On the relation of the entanglement spectrum to the bulk polarization

### 2021-02-03 version2/draft_mh.tex
small changes in the outlook

\title{On the relation of the entanglement spectrum to the bulk polarization \\ DRAFT}
\author{Carlos Ortega-Taberner}
\affiliation{Department of Physics, Stockholm University, AlbaNova University Center, SE-106 91 Stockholm, Sweden}
\affiliation{Nordita, KTH Royal Institute of Technology and Stockholm University, SE-106 91 Stockholm, Sweden}

@@

It was previously known that the bulk polarization (Zak phase, or geometric phase for $U(1)$ flux insertion) was encoded in the many-body ES. Here we have shown how it can be obtained from the single-particle entanglement spectrum of a non-interacting gapped system, showing that there is substantially more information in the single-particle ES than previously realized. This relation is simpler than the one involving the many-body ES, which is much more difficult to compute, and it provides more insight into the ES and EOS. In particular in the topological case, the relation between the quantized bulk polarization (or the Chern number in 2D) and the number of virtual topological edge states is demonstrated in a particularly transparent way. It also provides a new simple method for computing the bulk polarization for systems without translational invariance.

>>>>

\title{On the relation of the entanglement spectrum to the bulk polarization}
\author{Carlos Ortega-Taberner}
\affiliation{Department of Physics, Stockholm University, AlbaNova University Center, SE-106 91 Stockholm, Sweden}
\affiliation{Nordita, KTH Royal Institute of Technology and Stockholm University, SE-106 91 Stockholm, Sweden}

@@

It was previously known that the bulk polarization (Zak phase, or geometric phase for $U(1)$ flux insertion) was encoded in the many-body ES. Here we have shown how it can be obtained from the single-particle entanglement spectrum of a non-interacting gapped system, showing that there is substantially more information in the single-particle ES than previously realized. 
Our formulation is both simpler than the one involving the many-body ES, which is much more difficult to compute, and it provides additional insight into the ES and EOS. In particular in the topological case, the relation between the quantized bulk polarization (or the Chern number in 2D) and the number of virtual topological edge states is demonstrated in a particularly transparent way. It also provides a new simple method for computing the bulk polarization for systems without translational invariance.
